### 2005-11-15 doc/EFA/efa.tex
*** empty log message ***

\subsection{Background}
\newpage
\section{Finite Automaton (FA) with hierarchy}

\section{Extended Finite Automaton (EFA)}
The regular deterministic finite automaton $A$ is a $5$-tuple
\begin{eqnarray}
A=\langle Q,\Sigma,\delta, q_i,Q_m \rangle\nonumber
\end{eqnarray} where $Q$ is a finite set of states; $\Sigma$ (the alphabet) is a nonempty
set of events; $\delta$ is the transition function; $q_i$ is the
initial state and $Q_m \subseteq Q$ is the set of marked states.
An Extended Deterministic Finite Automaton (EFA) is an
augmentation of the regular finite automaton with guard and action
formulas associated to the transitions. The transitions in the EFA
are enabled if and only if a guard formula is true. Moreover, when
a transition in an EFA is taken, updating actions of a set of
finite integer variables may follow. Formally this is done as
follows.

@@

Formally we define the synchronous product $A||B$ of two EFA's
\begin{eqnarray}
A=\langle Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X^{A},
G^A,x^A_0,X^A_m \rangle
\end{eqnarray} and \begin{eqnarray} B=\langle Q^B, \Sigma^B, \delta^B, q^B_i,
Q^B_m, X^{B}, G^B,x^B_0,X^B_m \rangle
\end{eqnarray} as
\begin{eqnarray}
A||B&=&\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,
\delta^{A||B}, (q^A_i,q^B_i) ,\nonumber\\
&& Q^A_m\times Q^B_m, X^A \times X^B, G^{A||B},(x^A_0,x^B_0),
X^A_m \times X^B_m \rangle.
\end{eqnarray}
To define the guard functions, the action functions and the
transition function of the product we make use of the
corresponding functions in $A$ and $B$. To denote the elements of
$A||B$ we remove unnecessary  superscripts and write $A||B=\langle
Q, \Sigma, \delta, q_i, Q_m, X, G ,x_0, X_m \rangle$. The guard
functions $g_{q \sigma}(x)$ are defined as
\begin{eqnarray}\label{guard_sync_1}
g_{q \sigma}(x)=\left\{
\begin{array}{ll}
g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
g_{q^A\sigma}(x^{A})\quad\quad\quad\quad\quad\quad\quad\sigma\in \Sigma^A\setminus\Sigma^B\\
g_{q^B\sigma}(x^{B})\quad\quad\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
\end{array}\right.
\end{eqnarray}
If any one of the guards functions in the above definition are
undefined the guard function for the product $g_{q \sigma}$ is
also undefined. To simplify the notation further, we use "dummy"
guard functions in the following way. If $\sigma\in \Sigma^A
\setminus \Sigma^B$ we introduce fictional guards functions
$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. Doing this, we can write
equation (\ref{guard_sync_1}) as
\begin{eqnarray}
g_{q \sigma}(x)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
\end{eqnarray}
The updating functions $a_{q \sigma}(x)$ are defined as
\begin{eqnarray}\label{action_sync_1}
a_{q \sigma}(x)=\left\{
\begin{array}{ll}
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
\big(a^A_{q^A\sigma}(x^{A}),x^{B}\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^A\setminus\Sigma^B\\
\big(x^{A},a^B_{q^B\sigma}(x^{B})\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
\end{array}\right.
\end{eqnarray}
As in the definition of the guard functions the notion of being
undefined is inherited from the action functions of $A$ and $B$.
Similar to the "dummy" guard functions we introduce fictional
action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
$a_{q^A\sigma}(x^A)=x^A$. With this convention equation
(\ref{action_sync_1}) simplifies to
\begin{eqnarray}
a_{q \sigma}(x)=
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
\end{eqnarray}

@@

Given an extended finite automaton $A=\langle Q, \Sigma, \delta,
q_i, Q_m, X, G ,x_0, X_m \rangle$ with simple guard and action
functions, the algorithm can be divided into four steps:\\
- Form a regular FA, $A'$, from $A$ by removing $G$, $X$, $X_m$,
$x_0$ and change the names of all enabled events $\sigma$ to
$q\sigma$ (associate with the corresponding
state). Divide the event(transition) $q\sigma$ further into $|g_{q \sigma}|$ identical copies $\{(q\sigma)_k\}^{|g_{q \sigma}|}_{k=1}$.\\
- Realize an automata $A_{x_k}$ for each variable $x_k$, with
marked "states" $X_{mk}$,
 initial state $x_{k 0}$ and with as many events the total number of events of type $q\sigma$ in $A'$.\\
- Synchronize all automata $A_{x_k}$, $k=1,...,n$ and $A'$.\\
- Change back to original event names.
\newpage
\subsection{Synchronous composition of EFA's with common variables}

>>>>

\subsection{Background}
\newpage
\section{Finite Automata (FA) with hierarchy}

\section{Extended Finite Automata (EFA)}
The regular deterministic finite automaton $A$ is a $5$-tuple
\begin{eqnarray}
A=\langle Q,\Sigma,\delta, q_i,Q_m \rangle\nonumber
\end{eqnarray} where $Q$ is a finite set of states; $\Sigma$ (the alphabet) is a nonempty
set of events; $\delta$ is the transition function; $q_i$ is the
initial state and $Q_m \subseteq Q$ is the set of marked states.
An Extended Deterministic Finite Automaton (EFA) is an
augmentation of the regular finite automaton with guard and action
formulas associated to the transitions. The transitions in the EFA
are enabled if and only if a guard formula is true. Moreover, when
a transition in an EFA is taken, updating actions of a set of
finite integer variables may follow. Formally this is done as
follows.

@@

Formally we define the synchronous product $A||B$ of two EFA's
\begin{eqnarray}
A=\langle Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X^{A},
G^A,x^A_0,X^A_m \rangle
\end{eqnarray} and \begin{eqnarray} B=\langle Q^B, \Sigma^B, \delta^B, q^B_i,
Q^B_m, X^{B}, G^B,x^B_0,X^B_m \rangle
\end{eqnarray} as
\begin{eqnarray}
A||B&=&\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,
\delta^{A||B}, (q^A_i,q^B_i) ,\nonumber\\
&& Q^A_m\times Q^B_m, X^A \times X^B, G^{A||B},(x^A_0,x^B_0),
X^A_m \times X^B_m \rangle.
\end{eqnarray}
To define the guard functions, the action functions and the
transition function of the product we make use of the
corresponding functions in $A$ and $B$. To denote the elements of
$A||B$ we remove unnecessary  superscripts and write $A||B=\langle
Q, \Sigma, \delta, q_i, Q_m, X, G ,x_0, X_m \rangle$. The guard
functions $g_{q \sigma}(x)$ are defined as
\begin{eqnarray}\label{guard_sync_1}
g_{q \sigma}(x)=\left\{
\begin{array}{ll}
g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
g_{q^A\sigma}(x^{A})\quad\quad\quad\quad\quad\quad\quad\sigma\in \Sigma^A\setminus\Sigma^B\\
g_{q^B\sigma}(x^{B})\quad\quad\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
\end{array}\right.
\end{eqnarray}
If any one of the guards functions in the above definition are
undefined the guard function for the product $g_{q \sigma}$ is
also undefined. To simplify the notation further, we use "dummy"
guard functions in the following way. If $\sigma\in \Sigma^A
\setminus \Sigma^B$ we introduce fictional guards functions
$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. Doing this, we can write
equation (\ref{guard_sync_1}) as
\begin{eqnarray}
g_{q \sigma}(x)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
\end{eqnarray}
The updating functions $a_{q \sigma}(x)$ are defined as
\begin{eqnarray}\label{action_sync_1}
a_{q \sigma}(x)=\left\{
\begin{array}{ll}
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
\big(a^A_{q^A\sigma}(x^{A}),x^{B}\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^A\setminus\Sigma^B\\
\big(x^{A},a^B_{q^B\sigma}(x^{B})\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
\end{array}\right.
\end{eqnarray}
As in the definition of the guard functions (\ref{guard_sync_1}),
the notion of being undefined is inherited from the action
functions of $A$ and $B$. Similar to the "dummy" guard functions
we introduce fictional action functions. If $\sigma\in \Sigma^A
\setminus \Sigma^B$ we introduce fictional action functions
$a_{q^B\sigma}(x^B)=x^B$ for $B$ and if
$\sigma\in\Sigma^B\setminus \Sigma^A$ we set
$a_{q^A\sigma}(x^A)=x^A$. With this convention equation
(\ref{action_sync_1}) simplifies to
\begin{eqnarray}
a_{q \sigma}(x)=
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
\end{eqnarray}

@@

Given an extended finite automaton $A=\langle Q, \Sigma, \delta,
q_i, Q_m, X, G ,x_0, X_m \rangle$ with simple guard and action
functions, the algorithm can be divided into four steps:\\
- Form a regular FA, $A'$, from $A$ by removing $G$, $X$, $X_m$,
$x_0$ and change the names of all enabled events $\sigma$ to
$q\sigma$ i.e. the associate them with the corresponding
transition.\\
- Rename the events $q\sigma$ in $A'$ further into $|g_{q \sigma}|$ identical transitions with event names: $\{(q\sigma)_k\}^{|g_{q \sigma}|}_{k=1}$.\\
- Realize an automata $A_{x_k}$ for each variable $x_k$, with
marked "states" $X_{mk}$, initial state $x_{k 0}$ and with as many events the total number of events in $A'$.
The transitions correspond to the action functions and they are enabled if the guard formula is true.\\
- Synchronize all automata $A_{x_k}$, $k=1,...,n$ and $A'$.\\
- Change back to original event names.
\newpage
\subsection{Synchronous composition of EFA's with common variables}
